\documentclass{article}

\usepackage{amsmath}

\title{SI Model Stability}
\author{Charlotte Maryjean White}

\begin{document}
	\maketitle
	
	\section{Model}
	The SI (Susceptible-Infected) model  is a discrete-time, deterministic model.
	
	\[ \frac{dS}{dt}=-\frac{\alpha}{N}SI \]
	\[ \frac{dI}{dt}=\frac{\alpha}{N}SI \]
	
	\subsection{Parameters}
	
	\begin{itemize}
		\item $\alpha$: the contact rate, i.e. the average number of individuals successfully infected
	\end{itemize}
	
	$SI$ is the number of potential contacts with an infected individual. Then $\frac{\alpha}{N}SI$ is the average number of successful infections per population times the number of potentially successful contacts.
	
	\subsection{Initial Conditions}
	
	\[ S_0 + I_0 = N \]
	\[ S_0 > 0, I_0 > 0 \]
	
	\subsection{Assumptions}
	
	\begin{enumerate}
		\item The total population size, $ N $, remains constant: $ N = S_n + I_n $.
		\item Each individual is equally likely to contract the disease.
	\end{enumerate}
	
	\section{Stability Analysis}
	
	Reduce to one equation. Since $ S=N-I $:
	
	\[ \frac{dI}{dt}=\frac{\alpha}{N}(N-I)I \]
	
	The equilibria / fixed points are when $ \frac{dI}{dt}=0 $.
	\[ I^*=0, I^*=N \]
	
	Corresponding to $ S^*=N $ and $ S^*=0 $, respectively.
	
	This gives two $ (S,I) $ pairs: the DFE pair $ (N,0) $ and the endemic equilibrium pair $ (0,N) $.
	
	The Jacobian for $ x=\dfrac{dS}{dt} $ and $ y=\dfrac{dI}{dt} $ is
	
	\[
	J =
	\begin{vmatrix}
		\dfrac{\partial x}{\partial S} & \dfrac{\partial x}{\partial I} \\[6pt]
		\dfrac{\partial y}{\partial S} & \dfrac{\partial y}{\partial I}
	\end{vmatrix} =
	\begin{vmatrix}
		-\frac{\alpha}{N}I & -\frac{\alpha}{N}S \\[6pt]
		\frac{\alpha}{N}I & \frac{\alpha}{N}S
	\end{vmatrix} = \frac{\alpha}{N}
	\begin{vmatrix}
		-I & -S \\
		I & S
	\end{vmatrix}
	\]
	
	\subsection{Disease-Free Equilibrium (DFE)}
	
	DFE: $ (S,I)=(N,0) $
	
	Evaluate the Jacobian.
	\[
	J =
	\begin{vmatrix}
		0 & -\alpha \\
		0 & \alpha
	\end{vmatrix}
	\]
	
	Evaluate the characteristic polynomial:
	\[
	J - \lambda I =
	\begin{vmatrix}
		0 & -\alpha \\
		0 & \alpha
	\end{vmatrix} -
	\begin{vmatrix}
		\lambda & 0 \\
		0 & \lambda
	\end{vmatrix} = 
	\begin{vmatrix}
		-\lambda & -\alpha \\
		0 & \alpha - \lambda
	\end{vmatrix}
	\]
	
	Evaluate the determinant:
	\[
	det = -\lambda(\alpha - \lambda) - 0(-\alpha) = -\lambda(\alpha - \lambda)
	\]
	
	Eigenvalues: $ \lambda_1 = 0 $, $ \lambda_2 = \alpha > 0 $
	
	\subsection{Endemic Equilibrium}
	
	Endemic equilibrium: $ (S,I)=(0,N) $
	
	Evaluate the Jacobian.
	\[
	J =
	\begin{vmatrix}
		-\alpha & 0 \\
		\alpha & 0
	\end{vmatrix}
	\]
	
	Evaluate the characteristic polynomial:
	\[
	J - \lambda I =
	\begin{vmatrix}
		-\alpha & 0 \\
		\alpha & 0
	\end{vmatrix} -
	\begin{vmatrix}
		\lambda & 0 \\
		0 & \lambda
	\end{vmatrix} = 
	\begin{vmatrix}
		-\alpha - \lambda & 0 \\
		\alpha & - \lambda
	\end{vmatrix}
	\]
	
	Evaluate the determinant:
	\[
	det = -\lambda (-\alpha - \lambda) - 0(\alpha) =
	\lambda (\alpha + \lambda)
	\]
	
	Eigenvalues: $ \lambda_1 = 0 $, $ \lambda_2 = -\alpha < 0 $.
	
	\begin{thebibliography}{00}
		
		\bibitem{Allen1994} Allen, L.J.S. (1994). Some discrete-time SI, SIR, and SIS epidemic models. Mathematical Biosciences 124(1), 83-105. https://doi.org/10.1016/0025-5564(94)90025-6
		\bibitem{JacquezSimon1993} Jacquez, J.A. \& Simon, C.P. (1993). The stochastic SI model with recruitment and deaths I. comparison with the closed SIS model. Mathematical Biosciences, 117(1-2), 77-125. https://doi.org/10.1016/0025-5564(93)90018-6
		
	\end{thebibliography}
	
\end{document}